---
title: "Nasal asssimilation analysis - English monolingual"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

results_folder <- '/Users/marcofonseca/Documents/PhonRNN/EPhonRNN/results/'

library(ggplot2)
library(dplyr)
library(ggpubr)
library(ggfortify)
library(rstatix)
library(nortest)
library(tseries)
library(qqplotr)
library(DescTools)
library(FSA)
library(PMCMRplus)
library(AICcmodavg)
library(emmeans)



knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir='/Users/marcofonseca/Documents/PhonRNN/EPhonRNN/analysis/')
```
Opening and files and subsetting them.
```{r}
eval_data = read.csv(paste0(results_folder, "nasal_eval.csv"))
str(eval_data)

eval_data= eval_data[- grep(18, eval_data$following),] #deleting palatal nasals bc I could not deal with them 
eval_data= eval_data[- grep(20, eval_data$nasal),] #deleting uvular nasals bc I could not deal with them 
eval_data= eval_data[- grep(3, eval_data$nasal),] #deleting palatal nasals bc I could not deal with them 


#changing numbers to more meaningful names
eval_data["v_or_c"][eval_data["v_or_c"] == 0] <- "cons."
eval_data["v_or_c"][eval_data["v_or_c"] == 5] <- "vowel"


eval_data["nasal"][eval_data["nasal"] == 1] <- "m"
eval_data["nasal"][eval_data["nasal"] == 2] <- "n"
eval_data["nasal"][eval_data["nasal"] == 4] <- "ŋ"
eval_data["nasal"][eval_data["nasal"] == 6] <- "a"
eval_data["nasal"][eval_data["nasal"] == 7] <- "e"
eval_data["nasal"][eval_data["nasal"] == 8] <- "i"
eval_data["nasal"][eval_data["nasal"] == 9] <- "o"
eval_data["nasal"][eval_data["nasal"] == 10] <- "ɯ"

eval_data["following"][eval_data["following"] == 11] <- "ã"
eval_data["following"][eval_data["following"] == 12] <- "ĩ"
eval_data["following"][eval_data["following"] == 13] <- "ẽ"
eval_data["following"][eval_data["following"] == 14] <- "õ"
eval_data["following"][eval_data["following"] == 15] <- "ũ"

eval_data["following"][eval_data["following"] == 16] <- "labial"
eval_data["following"][eval_data["following"] == 17] <- "coronal"
eval_data["following"][eval_data["following"] == 19] <- "velar"





eval_data$following <- as.factor(eval_data$following)  # Convert character column to factor
eval_data$nasal <- as.factor(eval_data$nasal)  # Convert character column to factor
str(eval_data)


cons <- eval_data[eval_data$v_or_c == "cons.", ]
vs <- eval_data[eval_data$v_or_c == "vowel", ]


```
Plotting data 

```{r}
ggplot(cons) +
  geom_boxplot(aes(x=nasal, y=loss_mean, color=following), notch=TRUE)+
  xlab("Nasal consonant")+
  ylab("Average log  likelihood")+
   ggtitle("Average log likelihoods per nasal consonant and following sound\n Word set: test (13396 nonsense words)")+
  scale_color_brewer(name = "Following sound", palette="Dark2") 
  
ggplot(vs) +
  geom_boxplot(aes(x=nasal, y=loss_mean, color=following),  notch=TRUE)+
  xlab("Vowel")+
  ylab("Average log  likelihood")+
   ggtitle("Average log likelihoods per vowel and following sound\n Word set: test (24434 nonsense words)")+
  scale_color_brewer(name = "Following sound", palette="Dark2") 
  
```

Descriptive stats. 
```{r}

cons  %>%
  group_by(nasal, following) %>%
  get_summary_stats(loss_mean, type = "mean_sd")

vs  %>%
  group_by(nasal, following) %>%
  get_summary_stats(loss_mean, type = "mean_sd")


```
Another way of showing descriptive stats. 
n – The number of observations for each treatment.
mean – The mean value for each treatment.
sd – The standard deviation of each treatment.
stderr – The standard error of each treatment.  That is the standard deviation / sqrt (n).
LCL, UCL – The upper and lower confidence intervals of the mean.  That is to say, you can be 95% certain that the true mean falls between the lower and upper values specified for each treatment group assuming a normal distribution. 
median – The median value for each treatment.
min, max – The minimum and maximum value for each treatment. 
IQR – The inner quartile range of each treatment. That is the 75th percentile –  25th percentile.
LCLmed, UCLmed – The 95% confidence interval for the median.
```{r}
cons %>% select(nasal, following, loss_mean) %>% group_by(nasal, following) %>% 
  summarise(n=n(), 
            mean=mean(loss_mean, na.rm = FALSE), 
            sd=sd(loss_mean, na.rm = FALSE),
            stderr=sd/sqrt(n),
            LCL = mean - qt(1 - (0.05 / 2), n - 1) * stderr,
            UCL = mean + qt(1 - (0.05 / 2), n - 1) * stderr,
            median=median(loss_mean, na.rm = FALSE),
            min=min(loss_mean, na.rm = FALSE), 
            max=max(loss_mean, na.rm = FALSE),
            IQR=IQR(loss_mean, na.rm = FALSE),
            LCLmed = MedianCI(loss_mean, na.rm=FALSE)[2],
            UCLmed = MedianCI(loss_mean, na.rm=FALSE)[3])

vs %>% select(nasal, following, loss_mean) %>% group_by(nasal, following) %>% 
  summarise(n=n(), 
            mean=mean(loss_mean, na.rm = FALSE), 
            sd=sd(loss_mean, na.rm = FALSE),
            stderr=sd/sqrt(n),
            LCL = mean - qt(1 - (0.05 / 2), n - 1) * stderr,
            UCL = mean + qt(1 - (0.05 / 2), n - 1) * stderr,
            median=median(loss_mean, na.rm = FALSE),
            min=min(loss_mean, na.rm = FALSE), 
            max=max(loss_mean, na.rm = FALSE),
            IQR=IQR(loss_mean, na.rm = FALSE),
            LCLmed = MedianCI(loss_mean, na.rm=FALSE)[2],
            UCLmed = MedianCI(loss_mean, na.rm=FALSE)[3])
```



Checking for outliers: there is only one, I will keep it since its low likelihood is expected ( n+ velar). 

```{r}
outs <- eval_data %>% 
  group_by(following) %>%
  identify_outliers(loss_mean)
```

Inferential stats.
```{r}

model1 <- lm(log(-1*loss_mean) ~ nasal*following, data = cons)
model2 <- lm(log(-1*loss_mean)  ~ nasal+following, data = cons)


model.set <- list(model1, model2)
model.names <- c("int", "2way")

aictab(model.set, model.names)
autoplot(model1)

summary(model1)
emmeans(model1, list(pairwise~nasal*following))

model3 <- lm(log(-1*loss_mean) ~ nasal*following, data = vs)
model4 <- lm(log(-1*loss_mean) ~ nasal+following, data = vs)
model.set <- list(model3, model4)
model.names <- c("int", "2way")
aictab(model.set, model.names)
autoplot(model3)
summary(model3)
emmeans(model3, list(pairwise~nasal*following))


```




